Multi-target detection with application to cryo-electron microscopy

### 2019-03-04 inverse_problems/inverse_problems.tex
section 4.2 + proof

For the Poisson model: 
From these values, one may solve for the terms $\gamma a_x^1$, $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$, and $\gamma a_x^3[\ell_1,\ell_2]$. Indeed, recovering  $\gamma a_x^1$ and $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$ is immediate from \eqref{eq:mean_micrograph2} and \eqref{eq:ac2_micrograph2}, and $\gamma a_x^3[\ell_1,\ell_2]$ then follows from
%
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

The proof is provided in Section~\ref{sec:proof_uniqueness_poisson}.
\begin{proposition} \label{prop:uniqueness_poisson}
	Poisson result 
\end{proposition}

@@

\TODO{Check and specify if this holds for both models}
\TODO{Did we define densities $\gamma_k$?}
\TODO{We write the section for the Poisson model, but it would be the same for the Poisson model}

@@

\subsection{Proof of Proposition~\ref{prop:uniqueness_poisson}} \label{sec:proof_uniqueness_poisson}

	It is now easy to show how to solve for both $\gamma$ and $\sigma$ from these expressions, assuming $x$ is generic. Indeed, we observe the product (where $x[j]=0$ whenever $j < 0$ or $j > L-1$):
	%
	\begin{align}
	L^2(\gamma a_x^1) (\gamma a_x^2[1])
	&= \gamma^2\left( \sum_{i} x[i] \right) \left(\sum_{j} x[j]x[j+1] \right)
	\nonumber \\
	&= \gamma^2\sum_{j} \sum_{i} x[i]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{j} \sum_{\ell} x[j+\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell \ge 0} \sum_{j} x[j+\ell]x[j]x[j+1] 
	+ \gamma^2\sum_{\ell < 0} \sum_{j} x[j+\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell=0}^{L-1} a_x^3[1,\ell] 
	+ \gamma^2\sum_{\ell > 0} \sum_{j} x[j-\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell=0}^{L-1} a_x^3[1,\ell] 
	+ \gamma^2\sum_{\ell > 0} \sum_{j} x[j]x[j+\ell]x[j+\ell+1]
	\nonumber \\
	&= \gamma\left\{ \sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell]
	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]  \right\}.
	\end{align}
	%
	Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, we can form the ratio and solve for $\gamma$, and then $\sigma^2$:

	\begin{align}
	\gamma = L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
	{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
		+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]},
	\end{align}
	%
	\begin{align}
	\sigma^2 = a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
	- (2 L - 1) (a_y^1)^2.
	\end{align}

\section*{Acknowledgment} \TODO{To check}
The authors  thank Ayelet Heimowitz, Joe Kileel, Ti-Yen Kan and Amit Moscovich  for helpful discussions..
The research was partially supported by Award Number R01GM090200 from the NIGMS, FA9550-17-1-0291 from AFOSR, Simons Foundation Math+X Investigator Award, and the Moore Foundation Data-Driven Discovery Investigator Award.
NB is partially supported by NSF award DMS-1719558.

>>>>

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely. Furthermore, these quantities can be computed explicitly.

We first note that $\gamma a_x^1$, $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$, and $\gamma a_x^3[\ell_1,\ell_2]$ can be computed directly form the observed autocorrelations~\eqref{eq:mean_micrograph2},~\eqref{eq:ac2_micrograph2} and~\eqref{eq:ac3_micrograph2}.
 Indeed, recovering  $\gamma a_x^1$ and $\gamma a_x^2[\ell] + \sigma^2 \ell[\ell]$ is immediate from \eqref{eq:mean_micrograph2} and \eqref{eq:ac2_micrograph2}, and $\gamma a_x^3[\ell_1,\ell_2]$ then follows from
%
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}
Using these expressions, we prove the following in Section~\ref{sec:proof_uniqueness_poisson}.
\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, and the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely. 
In particular, 
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (\gamma a_x^2[1])}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2.
\end{align}
\end{proposition}

@@

\TODO{Check and specify if this holds for both models}
\TODO{Did we define densities $\gamma_k$?}
\TODO{We write the section for the well-separated model, but it would be the same for the Poisson model}

@@

\subsection{Proof of Proposition~\ref{prop:uniqueness_poisson}} \label{sec:proof_uniqueness_poisson}
	Let us assume that $x$ is generic. Indeed, we observe the product:
	%
	\begin{align}
	L^2(\gamma a_x^1) (\gamma a_x^2[1])
	&= \gamma^2\left( \sum_{i} x[i] \right) \left(\sum_{j} x[j]x[j+1] \right)
	\nonumber \\
	&= \gamma^2\sum_{j} \sum_{i} x[i]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{j} \sum_{\ell} x[j+\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell \ge 0} \sum_{j} x[j+\ell]x[j]x[j+1] 
	+ \gamma^2\sum_{\ell < 0} \sum_{j} x[j+\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell=0}^{L-1} a_x^3[1,\ell] 
	+ \gamma^2\sum_{\ell > 0} \sum_{j} x[j-\ell]x[j]x[j+1]
	\nonumber \\
	&=  \gamma^2\sum_{\ell=0}^{L-1} a_x^3[1,\ell] 
	+ \gamma^2\sum_{\ell > 0} \sum_{j} x[j]x[j+\ell]x[j+\ell+1]
	\nonumber \\
	&= \gamma\left\{ \sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell]
	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]  \right\}.
	\end{align}
	%
	Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, we can form the ratio and solve for $\gamma$, and then $\sigma^2$:

	\begin{align}
	\gamma = L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
	{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
		+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]},
	\end{align}
	%
	\begin{align}
	\sigma^2 = a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
	- (2 L - 1) (a_y^1)^2.
	\end{align}
Once $\gamma$ and $\sigma$ were computed, one can recover $x$ from $a_x^2$ and $a_x^3$  by Proposition~\ref{prop:uniqueness}.

\section*{Acknowledgment} \TODO{To check}
The authors  thank Ayelet Heimowitz, Joe Kileel, Ti-Yen Lan and Amit Moscovich  for helpful discussions..
The research was partially supported by Award Number R01GM090200 from the NIGMS, FA9550-17-1-0291 from AFOSR, Simons Foundation Math+X Investigator Award, and the Moore Foundation Data-Driven Discovery Investigator Award.
NB is partially supported by NSF award DMS-1719558.
